\chapter{Introduction}\label{chapter-01}
Historically, structural systems like buildings, bridges, and towers were conceived as fixed objects intended to remain in a single configuration. The geometry, connectivity, and primary function of these structural systems were established during the design phase and expected to remain unchanged throughout their service life. This design philosophy is closely aligned with the traditional priorities of structural engineering: durability, safety, and strength. Before the rise of modern engineering, resilience in structural systems was often achieved through traditional practices such as redundancy, oversizing, and reparability; by the nineteenth century, however, these tacit approaches were increasingly replaced by calculation-based design aimed at achieving safety with greater material efficiency~\cite{hassler2014resilience}. Coupled with parallel advances in analysis and construction, innovations in materials expanded the range of structural forms that could be realized and improved how buildings could perform, reflecting the long-standing reciprocal relationship between materials and architecture~\cite{bechthold2017materials}. Yet these advances largely refined structures intended to operate in one prescribed geometry and serve one primary function throughout their lives. Today, however, a growing range of applications demands structures that can do more than occupy a single configuration. Emergency infrastructure, rapidly deployable systems, compactly stowable structures, adaptive load-bearing assemblies, and future space-based construction all benefit from systems that can change shape or mechanical response in a controlled manner. The challenge, therefore, is no longer to design structures that stand efficiently, but to develop structural systems that can reconfigure without giving up structural usefulness.

The following sections of this chapter are organized as follows. Section~\ref{01-sec-reconfigurable-systems} will give a brief overview of the current state of reconfigurable systems. Section~\ref{01-sec-problem-statement} narrows the scope of this work and identifies the issues that prevent their direct adoption in large-scale structural systems. Section~\ref{01-sec-objectives} states the key objectives of this dissertation. Section~\ref{01-sec-organization} summarizes the highlights of the dissertation that will be presented in the following chapters.

\section{Current State of Reconfigurable Systems in Engineering}\label{01-sec-reconfigurable-systems}
Over the past century, researchers have developed a wide range of deployable and reconfigurable systems for applications that demand compact storage for efficient transportation, rapid deployment from stowed configuration, and controlled reconfiguration for functional use cases. Much of this work has emerged from automotive, aerospace, mechanical, and robotics engineering, where kinematic performance, packing efficiency, and actuation are primary design drivers. The resulting body of literature spans many distinct families of systems, including linkage-based mechanisms, compliant mechanisms, tensegrity assemblies, scissor and expanding geometries, origami-inspired structures, and pneumatic or membrane-based deployable structures. These structures have been studied extensively as a class of systems capable of transforming from a compact state to a larger functional configuration, and recent work continues to expand their geometric and analytical sophistication. However, their direct applicability to large-scale structural systems of the kind commonly encountered in civil engineering remains more limited and often highly application-specific. Even in recent large-scale origami work, a recurring challenge is that systems that fold well often do not carry large loads well, while systems that carry large loads often lose much of their deployability or adaptability. As a result, the literature offers many techniques for creating reconfigurable systems, but far fewer clear frameworks for introducing reconfigurability into existing static structural designs that engineers already trust. This gap motivates the central question of this dissertation: how can the benefits of reconfigurability be systematically embedded into conventional structural topologies without abandoning structural usefulness?

\section{Problem Statement}\label{01-sec-problem-statement}
The dissertation is about a systematic transformation from static structural topology to reconfigurable system to functional behavior.

Focus is on identifying ways to start with established structural topologies and transform them into systems capable of controlled reconfiguration while preserving meaningful structural behavior. Not on inventing new moving structures.

This dissertation asks how static structural topologies can be systematically transformed into reconfigurable systems that retain structural relevance while gaining controlled mobility and functional performance.

\section{Objectives of the Dissertation}\label{01-sec-objectives}
\begin{itemize}
\item Develop a systematic framework for transforming static truss topologies into reconfigurable structural systems.
\item Establish computational and mechanical tools for analyzing reconfigurable systems with joint compliance and large-displacement response.
\item Demonstrate how reconfigurable structural systems can be designed for functional behaviors such as deployment, energy absorption, and controlled motion.
\item Extend the broader static-to-reconfigurable viewpoint to shape-morphing structural forms, including curved-origami-inspired hull systems.
\end{itemize}

\section{Organization of the Dissertation}\label{01-sec-organization}
This dissertation investigates how static structural topologies can be transformed into functional reconfigurable systems. To address this objective, the dissertation is organized as follows:

\textbf{Chapter~\ref{chapter-02}} introduces a method for transforming static trusses into reconfigurable systems. This method uses Grahsof's flat-foldability criterion for planar quadrilateral linkages to strategically introduce a new node in every triangular unit of a given truss. These added nodes convert triangular units into flat-foldable quadrilateral linkages, thereby introducing kinematic degrees of freedom that enable global reconfigurability. A general workflow is presented for applying these transformations to most single-chorded truss designs, and it is supported by a sequential kinematic analysis framework to study the deployment behavior of the resulting systems. This workflow is first demonstrated on traditional truss designs and then extended to topology-optimized trusses with arbitrary geometries, to highlight its broader applicability. The chapter also presents experimental structural analysis of proof-of-concept prototypes to demonstrate fabrication feasibility and to show that the transformed systems retain load-bearing capacity and stiffness of the original designs.

\textbf{Chapter~\ref{chapter-03}} develops a Three-node Torsional Spring (3NTS) element to introduce rotational stiffness at the folding joints of the reconfigurable trusses developed in Chapter~\ref{chapter-02}. The chapter presents a first-principles mathematical formulation of a torsional coil spring element and its implementation within a nonlinear structural solver. The formulation is validated against analytical solutions for two benchmark examples, demonstrating its ability to capture structural instabilities associated with elastic buckling. The chapter then shows how torsional springs can be integrated into reconfigurable trusses to provide functional benefits such as resistance to load reversal, controlled path-dependent deployment, and energy absorption under external loading.

\textbf{Chapter~\ref{chapter-04}} builds on the 3NTS formulation from Chapter~\ref{chapter-03} to design bi- and multi-stable torsional springs. The reconfigurable truss systems augmented with these unique spring types enables the exploration of reconfigurable truss systems for function, including energy absorption, controlled deployment, and path-governed response.

\textbf{Chapter~\ref{chapter-05}} broadens the perspective of the dissertation by extending the transformation of static structural topology beyond civil engineering and into naval architecture. In this chapter, the focus shifts from reconfigurable truss systems to shape-morphing hull structures inspired by curved-crease origami. The chapter develops curved-crease-origami-inspired planing hulls that can be deployed rapidly from flat sheets, match the shape and therefore, hydrodynamics of conventional planing hulls, and utilize active folding to change shape and tune hydrodynamic performance of the hull on demand. A proof-of-concept prototype is also presented, together with a discussion of the steps required for real-world implementation.

\textbf{Chapter~\ref{chapter-06}} summarizes the main contributions of the dissertation, discusses its limitations, and outlines directions for future research.

Two appendices are included to provide additional background for the analytical tools and methods used throughout the dissertation.

\textbf{Appendix~\ref{01-appendix-seq-kinematics}} presents the mathematical formulation for the positional analysis of planar quadrilateral linkages, which serves as the basis of the sequential kinematic analysis framework used to study deployment of reconfigurable trusses. This appendix also includes the positional analysis of the special six-bar linkage case discussed in Chapter~\ref{chapter-02}, together with a note on interpreting the percentage of actuation of each kinematic degree of freedom in the reconfigurable truss systems.

\textbf{Appendix~\ref{02-appendix-truss-videos}} contains the supplementary videos of fabricated reconfigurable truss prototypes and their load-tests, to demonstrating failure type and load-carrying capacities, provided with Chapter~\ref{chapter-02}.

\textbf{Appendix~\ref{03-appendix-powersea}} describes the theoretical basis of the Powersea software used for the hydrodynamic analyses of planing hulls in Chapter~\ref{chapter-05}. This appendix briefly outlines the \textit{low-aspect-ratio strip theory} used in the software, states the key assumptions, and explains why this approach is suitable for comparative hydrodynamic studies of curved-crease origami and conventional planing hulls.