\chapter{Conclusions and Future Work}\label{chapter-06}
\section{Key Contributions of the Dissertation}\label{06-sec-contributions}
This work presents a systematic method for converting static trusses into reconfigurable, shape-morphing systems using principles of flat-foldable quadrilateral linkages. The key contributions are as follows:

\textit{General rules for node placement that enable reconfigurability in the triangular components of static trusses}. Our proposed method transforms triangular units of static trusses into quadrilateral linkages by introducing an additional node. The node placement is guided by closed-form expressions derived for all triangle geometries (Fig.~\ref{fig-node-intro}) to satisfy the Grashof flat-foldability criterion in the resulting quadrilateral linkages. These nodes are exclusively placed on tensile members to ensure the structure remains stable under external loads and carries its design load, even with mobility greater than one. Static trusses transformed using this method can be reconfigured to minimize their convex hull area and stored much compactly upon the actuation of their kinematic degrees of freedom.

\textit{A programmatic workflow to transform static trusses into reconfigurable systems and study their kinematic behavior.} We developed a workflow that extends unit-level transformations to the system level. This workflow guides the selection of the candidate members based on the geometry of the triangular units and their member forces in most statically determinate trusses with a continuous series of triangular units between the top and bottom chords. We demonstrate the applicability of our workflow with the help of three commonly observed truss designs (Scissor, Fink, and Warren) and trusses of arbitrary geometries obtained from topology optimization (cantilever, L-shaped, and Serpentine). Additionally, we developed a sequential kinematic simulator to visualize the motion path of the reconfigurable trusses and evaluate their packing density as they morph into different configurations upon sequential actuation of their kinematic degrees of freedom.

\textit{Fabricating proof-of-concept prototypes to demonstrate structural capacities of reconfigurable trusses.} We validated the feasibility of constructing real-life reconfigurable trusses using our method by fabricating two proof-of-concept prototypes \textemdash{} a 30~cm reconfigurable cantilever truss and a 3~m Warren truss. Both prototypes employ a joint design strategy that distributes truss members in and out of the plane to ensure desired shape-morphing behavior without member contact or overlap. Load testing of the cantilever trusses confirmed that the reconfigurable truss exhibits force-stable behavior and maintains a peak load capacity and structural stiffness comparable to its static counterparts despite having mobility greater than one. The three-meter-long reconfigurable Warren truss weighing 38~kg supported a peak load of 19~kN, demonstrating the structural capacity and scalability of the proposed designs.

Using reduced-order numerical simulations for curved origami folding and hydrodynamic analysis, we have explored the versatility of curved-crease origami hulls. In particular, the study has demonstrated the following advantages associated with curved-crease origami hulls.

\noindent{}\textit{Rapidly deployable hulls that start as flat sheets and conform to desired shapes}. The curved-crease origami hulls introduced in this study can be designed to conform to a desired hull shape by adjusting and optimizing the geometric parameters of the initial curved-crease pattern. Shape-matching analysis reveals that these hulls exhibit a mean Hausdorff distance of less than 0.5\% of the overall hull length across much of the planing surface that would interact with water. Consequently, traditional hull designs like the VPS (deep-V, Planing hull with straight surface) can be fabricated using curved-crease origami techniques. Despite being non-developable, the entire curved-crease origami hull can be fabricated in a flat-folded state. This flat-foldability presents advantages over conventional fabrication methods, such as facilitating off-site construction and efficient storage and transportation. Moreover, the curved-crease origami hulls can be deployed rapidly from their flat folded state. This attribute is particularly advantageous in scenarios such as post-disaster relief operations for search and rescue, as well as the deployment of multiple uncrewed aquatic vehicles for naval operations, where prompt response, multiple distributed vessels, and adaptability are paramount.

\noindent{}\textit{Replicating hydrodynamics of conventional hull forms}. By extension of the ability to conform to conventional hull shapes, the curved-crease origami hulls demonstrate hydrodynamic performance that closely matches their conventional counterparts (VPS and GPPH). The hydrodynamic simulations performed with Powersea in still water conditions reveal nearly equivalent trends in performance characteristics such as friction drag, total resistance, hull sinkage, and trim over a wide range of forward speeds. In wavy water (head seas) conditions, the curved-crease origami hulls demonstrate a notable similarity in the amplitude and frequency of the hull motion time histories to their conventional counterparts. These observations suggest that hulls designed with curved origami principles can be expected to deliver comparable performance to the traditional hulls, that is, maintain power efficiency, passenger comfort, and safety, among other performance parameters, in addition to the inherent advantages offered by origami fabrication techniques.

\noindent{}\textit{Shape morphing hulls for on-demand tunable hydrodynamics}. Hulls designed with curved origami principles grant control over the angle of the fold lines along the keel and chine of the hull. The extent of folding (or actuation) of these fold lines facilitates on-demand morphing of the hull shape, allowing for the modification of key geometric features, such as the deadrise angle. Notably, hulls configured with a lower deadrise exhibit a flatter planing surface, making them well-suited for calm water conditions. They typically maintain lower sinkage and trim angles, allowing them to experience lower resistance and thus achieve more efficient motion. However, this configuration experiences increased amplitudes of hull motion in waves, making it less suitable for effective seakeeping in wavy conditions. Conversely, the same hull configured with a higher deadrise angle features a sharp V-shaped planing surface. While this configuration typically encounters higher total resistance, sinkage, and trim angles in still waters, the amplitude of hull motion is significantly lower in wavy water conditions compared to their low deadrise counterparts. Thus, a singular curved-crease origami hull can transition between various deadrise configurations, rendering it suitable for use in various sea conditions. Furthermore, tuning its hydrodynamic performance also ensures increased passenger comfort and operational efficiency.

\section{Considerations for Future Work}\label{06-sec-future-work}
While the reconfigurable trusses introduced in this work offer exciting avenues for designing shape-morphing structures, several challenges must be overcome before translating these designs into practical, life-sized structures. Future research will focus on tackling the following challenges:

\textit{Mitigating structural overlap to optimize packing density.} As the complexity and degrees of freedom for reconfigurable trusses increase, the potential for structural self-overlapping during actuation becomes increasingly evident. This issue is particularly highlighted in the Scissor truss example illustrated in Fig.~\ref{fig-trad-truss}\,(a), where the actuation of the third degree of freedom results in portions of the structure overlapping. Real-life truss structures cannot achieve a compact state without providing special provisions for folding mechanisms. For such structures, it is crucial to understand their spatial movement patterns and identify the degrees of freedom whose actuation triggers a self-overlap. Addressing this challenge necessitates future investigation into varied sequences of degree-of-freedom actuation to mitigate structural overlap. Furthermore, automating the distribution of truss members out of the plane and incorporating additional methods for node placement (compressive members) will enhance the feasibility of reconfigurable structures. These practices can be explored to prevent self-overlapping and optimize the ability of the structure to minimize its footprint after shape transformations.

\textit{Handling multi-chord trusses, series of similar triangles and polygons in a bar-linked structure.} The node introduction workflow described in Section~\ref{02-subsec-algo} effectively transforms many commonly used truss structures into reconfigurable systems. However, the simplifying assumptions used herein limit its applicability when addressing the following three specialized cases. First, trusses with multiple layers of triangles between their top and bottom chords demand meticulous consideration during node introduction. Adding a new node in these configurations is challenging, as it influences the mobility of adjacent triangular units. If not executed carefully, this modification can lead to incompatible kinematics at the unit level, and structural instability and reduced packing efficiency at the system level. Second, truss structures composed of a sequence of similar triangles, as found in Pratt trusses, present a unique challenge while introducing nodes. These repeating patterns often cause certain triangles to be skipped by the current workflow. Therefore, further adjustments are necessary to ensure that nodes are added appropriately and that the structure achieves the desired reconfigurability. Lastly, trusses for arbitrary geometries derived from topology optimization often exhibit complex forms and include polygons with more than three sides. The current workflow is not equipped to handle such complex structures. Future work can focus on strategies that help simplify the structures and refine the workflow to address such complex cases effectively. Alternatively, topology optimization algorithms can be improved to provide optimal truss designs that can be directly transformed into reconfigurable systems. In addition, graph-theory techniques could also be explored to generalize the design process of reconfigurable trusses.

Addressing the aforementioned research topics will pave the way for the development of reconfigurable trusses that effectively balance structural capacity, stiffness, and stability while enabling shape morphability, facilitating on-demand deployment, and ensuring compact storage.

Several items remain unexplored, particularly concerning the material selection, actuation methods, the structural integrity of the hull, and an in-depth exploration of the interplay between origami hull shape morphing and hydrodynamics. Future work could focus on:

\noindent{}\textit{Optimizing material selection for curved panels and fold joints}. Curved-crease origami panels and fold joints experience large strains throughout their deployment process. In addition to the large strains, panels and creases of deployable structures are vulnerable to damage due to long-term storage and fatigue as they underdo repeated folding and unfolding~\cite{ma_design_2024}. In CCO hulls, fatigue-induced cracks can compromise the hull's watertightness, posing a risk of sinking.

To fabricate multi-use deployable curved-crease origami hulls, it is essential to identify flexible materials that can withstand numerous bending and folding cycles while demonstrating durability and resistance to the corrosive marine environment. Current deployable structures with highly deformable components like rigid-hull inflatable boats~\cite{halswell_rhib_2011} and pneumatic floats for inflatable bridges~\cite{russell_causeway_2014} use materials like Hypalon\textregistered{} coated polyester, PVC, or Polyurethane, which last for 15-20 years before showing signs of damage. Recent work has also highlighted the potential of flexible and durable connections for origami systems using carbon fiber-reinforced polymer and urethane composites~\cite{deleo_origami_2020, deleo_composite_2018}. While the proof-of-concept prototypes in Section~\ref{05-subsec-physical-implementation} present initial feasibility, future work must focus on: (\textit{i}\,) the selection of flexible composite materials that are resilient to fatigue damage and (\textit{ii}\,) advanced bonding to connect creases and panels. The ongoing work on CCO hull fabrication at NSWCCD~(Kelly and Hart, Personal Communication, 2024) also aims to address the issue of fatigue in origami creases by developing cast-in-place neoprene joints for the CCO panels.

\noindent{}\textit{Actuation methods and structural stiffeners for enhanced stability}. Future research should explore actuation methods for efficiently deploying curved crease origami hulls. Depending on the hull's material, size, and intended application, an investigation of the forces required to transform the hull from a flat-folded to a deployed state must be undertaken. Analysis of the required actuation forces will guide the selection of deployment methods for curved-crease origami hulls (for example, mechanized devices, inflatable systems, or deployment by hand).

The post-deployment strengthening of the CCO hulls is an equally essential area for future exploration. Although we have presented a proof-of-concept in Section~\ref{05-subsec-physical-implementation} for the inclusion of stiffeners, future work must focus on: (\textit{i}\,) incorporating deployable bulkheads, girders, or frames for post-deployment stiffening of the CCO hulls, (\textit{ii}\,) exploring connectors that ensure compatible kinematics between the CCO hull shell and deployable stiffeners and (\textit{iii}\,) studying the quantitative increase of structural stiffness provided by the stiffeners against shell collapse and torsional deformations.

\noindent{}\textit{Evaluating structural integrity for various payloads under wave impact loading}. A critical area of consideration for curved-crease origami hulls is the evaluation of their structural integrity when subject to dynamic wave impact loads. Planing hulls are known to experience high pressures locally in the bow area. It is, therefore, essential to assess the local strength of the hull shell, mainly when the thickness of the hull is limited by surface curvature (as shown in Fig.~\ref{fig-prototype}\,(a)). Thus, future work must study the effects of varying payloads on the pressures and forces experienced by an origami hull in different deadrise configurations. This study could be performed using tow tank tests or CFD hydrodynamic analysis to evaluate pressures and forces on the hull surface and subsequently using structural finite elements analysis to evaluate the structural response, including stresses and strains.

\noindent{}\textit{In-depth exploration of the relationship between shape morphing and hydrodynamics}. The shape morphing ability of curved-origami hulls not only influences the hull deadrise but also alters various geometric parameters such as the overall length, bow curvature, beam width, center of gravity, and radius of gyration. Our work thus far has highlighted the potential of the CCO hulls' shape-morphing ability for the tuning of hull hydrodynamics. However, we have yet to establish the relationship between the combination of these geometrical transformations and hull hydrodynamics. While Powersea is an excellent tool for analyzing planing hulls with minimal computational costs, a significant amount of valuable information is lost due to the underlying assumptions and simplifications of the flow. These simplifications include the reduction of the order from 3D to 2D and linearizing the free surface condition within each time step, among others. Thus, future work must focus on conducting one-way coupled fluid-structure interaction analyses to comprehensively understand the relationship between the CCO shape change and hull hydrodynamics. This relationship can explored in greater detail by examining the Response Amplification Operators (RAOs), such as the heave and pitch transfer functions, for the varying configurations of the CCO hulls subjected to different wavelengths. By analyzing RAOs, we can better understand how CCO shape changes tune parameters like the resonant frequency of the hull. Moreover, these analyses will simulate hull performance with high confidence levels, allowing us to evaluate characteristics such as added wave resistance, whisker spray resistance, and accelerations at various points on the hull to assess the sea-keeping fitness of CCO hulls. Additionally, tank tests may also be performed to support the CFD results.

Addressing the research topics discussed above will help lead to the future implementation of deployable curved-crease origami planing hulls with on-demand performance tunability and optimization of crew comfort for civil and military applications.